Fail Faster

### 2025-03-15 paper/paper.tex
minor updates

\subsubsection{Monadic DSL Abstraction Overhead}
\jwc{
  \begin{itemize}
    \item It's been long-known that clean functional abstractions have a runtime overhead (this should be familiar by this time in the paper).
    \item How does (simplified) BQ work?
    \begin{itemize}
      \item The basic generator type: \texttt{'a generator = int -> SR.t -> 'a}. Size and random seed to deterministic value. (note: \texttt{SR.t} is a mutable seed)
      \item This gets a monad intance in the obvious way (show code).
      \item Also show the code for \texttt{int}, how it calls the underlying SR function.
    \end{itemize}
    % \item Note that \emph{extensionally} \texttt{generate (create (fun size random -> e)) size random = e}, but the OCaml compiler does not always perform this optimization, or do the inlining required to expose it.
    % (When sufficiently obfuscated behind returns and binds ...) This program compiles to code that (1) allocates the closure for `e', (2) passes it to create (which returns the closure), and then calls (3) generate, which immediately jumps into the closure.
    \item Show benchmarks of the running example, versus the version where you inline everything. ()
    \item Let's look at the running example: inline it all the way.
  \end{itemize}
}

@@

\subsubsection{Combinator Abstraction Overhead}

@@

% \subsubsection{Function call overhead}
\subsubsection{Choice of RNG}
\jwc{We really need to have explained the effect ordering equivalence aspect of the evaluation before here.}

The core of any PBT generator library is a (pseudo) random number generator.
Different PBT libraries use a wide range of different RNGs.
Following the original Haskell QuickCheck, \bq uses the SplitMix algorithm \cn,
implemented (statefully) as an OCaml library called \texttt{Splittable\_random}. Meanwhile,
ScalaCheck uses the JSF algorithm \cn.
% https://www.pcg-random.org/posts/bob-jenkins-small-prng-passes-practrand.html
% https://burtleburtle.net/bob/rand/smallprng.html

@@

\section{Eliminating Abstraction Overhead of Generator DSLs by Staging}
\jwc{Emphasize that this is an \emph{COMPLETELY EQUIVALENT drop in replacement!} We din't just build a different library with different distributions.}

@@

\subsection{Fast Random Number Generators}
\label{subsection:faster-rng}

\jwc{
  \begin{itemize}
    \item Using the fastest RNG possible is important.
    \item In OCaml, the splittalbe random RNG is not as fast as it could be, because of the Ints issue.
    \item In this section, we describe how to resolve the ints issue by calling out ot the FFI. 
    \item All of this can be done with Jane Street bleeding-edge compiler extensions, though these are not yet compatible with MetaOCaml.
    \item Because different libraries 
  \end{itemize}
}

@@

\jwc{NOTE: we should test generator speed across both languages, but speed -> bugfinding ability in only OCaml.}
\jwc{
  Baseline generators to test speed in both languages:
  \begin{itemize}
    \item Single int
    \item Pair of ints, constrained
    \item List of ints
  \end{itemize}
}
\subsection{Impact on bug-finding ability}
\jwc{Staging type-derived generators (which are the most commonly used ones) can turn a timeout into a found bug!}
\section{Conclusion \& Future Work}
\section{Related Work}
\jwc{A generator is a parser of randomness, and people have implemented lots of staged parser libraries!}
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bib}

>>>>

\subsection{Monadic DSL Abstraction Overhead}
\jwc{
  \begin{itemize}
    \item It's been long-known that clean functional abstractions have a runtime overhead (this should be familiar by this time in the paper).
    \item How does (simplified) BQ work?
    \begin{itemize}
      \item The basic generator type: \texttt{'a generator = int -> SR.t -> 'a}. Size and random seed to deterministic value. (note: \texttt{SR.t} is a mutable seed)
      \item This gets a monad intance in the obvious way (show code).
      \item Also show the code for \texttt{int}, how it calls the underlying SR function.
    \end{itemize}
    % \item Note that \emph{extensionally} \texttt{generate (create (fun size random -> e)) size random = e}, but the OCaml compiler does not always perform this optimization, or do the inlining required to expose it.
    % (When sufficiently obfuscated behind returns and binds ...) This program compiles to code that (1) allocates the closure for `e', (2) passes it to create (which returns the closure), and then calls (3) generate, which immediately jumps into the closure.
    \item Show benchmarks of the running example, versus the version where you inline everything. ()
    \item Let's look at the running example: inline it all the way.
  \end{itemize}
}

@@

\subsection{Combinator Abstraction Overhead}

@@

% \subsubsection{Function call overhead}
\subsection{Choice of RNG}
\jwc{We really need to have explained the effect ordering equivalence aspect of the evaluation before here.}

The core of any PBT generator library is a random number generator: to generate random values, we need
Different PBT libraries use a wide range of different RNGs.
Following the original Haskell QuickCheck, \bq uses the SplitMix algorithm \cn,
implemented (statefully) as an OCaml library called \texttt{Splittable\_random}. Meanwhile,
ScalaCheck uses the JSF algorithm \cn.
% https://www.pcg-random.org/posts/bob-jenkins-small-prng-passes-practrand.html
% https://burtleburtle.net/bob/rand/smallprng.html

@@

\section{Eliminating Abstraction Overhead of Generator DSLs by Staging and Faster RNG}
\jwc{Emphasize that this is an \emph{COMPLETELY EQUIVALENT drop in replacement!} We din't just build a different library with different distributions.}

@@

\subsection{Faster SplitMix with Unboxed Int}
\label{subsection:faster-rng}

\jwc{Unclear what we should call this section}

\jwc{
  \begin{itemize}
    \item As we discussed in Section~\ref{section:motiv}, inefficient RNG choice is a bottleneck for bugfinding power.
    \item Faster RNGs and sources of entropy exist, though they are extensionally inequivalent. To demonstrate the bugfinding power of a faster RNG, we use the natural experiment provided by OCaml's inefficient implementation of SplitMix.
    \item Unlike the previous subsections, this is not a anguage-agnostic technique. In most other PBT frameworks, the RNG operates on by machine integers, so this inefficiency does not exist.
    \item We are simply using this as a way to demonstrate that faster RNG can translate to faster bugfinding for sampling-bounded generators.
    \item Here's how SplitMix is slow: the core of SplitMix is nextInt64, which does allocations for each operation.
    \item If we could use actual int64s, we wouldn't have to allocate.
    \item So we do this:
    \begin{itemize}
      \item We implement SplitMix in C (which uses normal Int64s) and real int64 arithmetic, and use the OCaml FFI to call out to the C functions. T
      \item This could be done with Jane Street bleeding-edge compiler extensions, though these are not yet implemented in a a version of the OCaml compiler that also has staging.
    \end{itemize}
  \end{itemize}
}

@@

\jwc{NOTE: we should test generator speed across both languages, but speed -> bugfinding ability in only OCaml.}
\jwc{
  Baseline generators to test speed in both languages:
  \begin{itemize}
    \item Single int
    \item Pair of ints, constrained
    \item List of ints without bind (use map): lots of sampling, minimal binds.
    \item Unableled Trees of a fixed size (no weighted union) minimal sampling, lots of binds.
  \end{itemize}
}
\subsection{Impact on bug-finding ability}
\jwc{Staging type-derived generators (which are the most commonly used ones) can turn a timeout into a found bug!}
\section{Conclusion \& Future Work}
\section{Related Work}
\jwc{A generator is a parser of randomness, and people have implemented lots of staged parser libraries!}
%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bib}
